---
title: "bike Share Analysis"
author: "Joe Tudor"
date: "15 Oct 2020"
output: html_notebook
---

This analysis is to better understand the bike share program data in Washington, D.C. and discover if there are relations that we can find to predict ridership.


Prepare workspace:

## Load packages

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(skimr)
library(summarytools)
library(lubridate)
```

## Read and clean the data
Loading the data
```{r}
dfa= read_excel("bikesharedata.xlsx", sheet = 2) %>%
clean_names()
```
Skimming the data for a rough look and the data set size, variables, and datatypes.
```{r}
skim(dfa)
```

getting rid of the unneeded ride_id just to unclutter the data a bit, also removes trips with a negative trip duration value.
```{r}
dfb = dfa %>% select(-ride_id) %>% filter(duration>0) 
```
plots the data with respect to duration and start time on a scatterplot.
```{r}
dfb %>% filter(mday(started_at)==3,duration<60) %>% ggplot(aes(x=started_at,y=duration))+
  geom_point(alpha=.1)+
  ggtitle("Trip duration vs start time (August 3)")
```




removing the observations that have an observation too high or negative that would indicate bad data
```{r}
dfc = dfb %>% filter(duration>0,duration<1440, )
```

plotting the data in a histogram to visualize the ride durations
```{r}
dfc %>% filter(duration<300) %>% ggplot(aes(duration)) + geom_histogram(binwidth = 1)
```

```{r}
dfsorted= dfc %>% dplyr::arrange(duration)
```
breaking down the duration histogram by weekdays
```{r}
dfsorted %>% filter(duration<300) %>% ggplot(aes(x = duration)) + geom_histogram(binwidth = 1) + facet_wrap(~weekday)
```
```{r}
dfsorted %>% filter(mday(started_at)==3,duration<60) %>% ggplot(aes(x=started_at,y=duration))+
  geom_point(alpha=.1)+
  ggtitle("Trip duration vs start time (August 3)")
```
How many bikes are being used at any given time of day?
To solve this problem we must look at the start and end times. First we slice the dataset to take a sample of 100.
```{r}
dfe=dfsorted %>% filter(mday(started_at)==3) %>% slice_sample(n=100)
```

```{r}
dfe %>% select(start=started_at,end=ended_at)
```
We pivot the data set so that each start and end time is recorded in the table as an observation in a sequence. 
```{r}
dfe %>% select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end) 
```
```{r}
dfe %>% select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t)
```
After adjusting the names of the new column variables we create a new column to increment the number of bikes at a given time. Each start time will produce an increment of 1 while each end time will produce an increment of -1 since the rider is no longer using the bike and can be removed from the current cumulative total.
```{r}
dfe %>% select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) 
```
We now simply create a column to represent the cumulative total that reflects the number of riders at any given time from this sample.
```{r}
dfe %>% select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  mutate(riders=cumsum(increment))
```
Next we look at the ridership comparing the docked vs electric bike as well as the ridership with respect to each day of the week.
```{r}
dfb %>% 
  filter(month(started_at)==8,month(ended_at)==8) %>%
 select(rideable_type,start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  group_by(rideable_type) %>%
  mutate(riders=cumsum(increment)) %>% filter(mday(t)<=7) %>%
    ggplot(aes(t,riders,color=rideable_type)) + geom_step() +
  facet_wrap(~mday(t),scales = "free_x",ncol = 7)
```
```{r}
dfr = dfb %>% 
  filter(month(started_at)==8,month(ended_at)==8) %>%
 select(rideable_type,start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  group_by(rideable_type) %>%
  mutate(riders=cumsum(increment)) 
```
This dataframe looks at the ridership over the month of august using the cumulative sum technique used earlier.
```{r}
dfr1 = dfb %>% 
  filter(month(started_at)==8,month(ended_at)==8) %>%
 select(start=started_at,end=ended_at) %>%
  pivot_longer(start:end, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start"~1,
   type=="end" ~ -1
  )) %>%
  mutate(riders=cumsum(increment)) 
```
We try to find the max riders in any hour time interval.
```{r}
dfrh = dfr1 %>% group_by(hour(t)) %>% summarise(average_riders = mean(riders))
```
We summarize the data to look at ridership with relation to week day, focusing on hour intervals for each of the days.
```{r}
dfr1 %>% group_by(hour(t),wday(t,label = TRUE)) %>%summarise(max_riders = max(riders),average_riders = mean(riders))
```

Challenge Problem:
usage at any particular station
start from a particular location

