---
title: "Class Enrollment Data:MA132"
author: "Joe Tudor"
date: "21 Sep 2020"
output: html_notebook
---

THe intention of this analysis is to try to predict next semester's enrollment(spring 2021) for MA132.


Prepare workspace:

## Load packages

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(skimr)
library(summarytools)
```

## Read and clean the data

```{r}
dfa= read_excel("CU_ACAD_DEPT_ENRL.xlsx") %>%
clean_names()
```

```{r}
skim(dfa)
```

### Tidy Data

lets seperate column Term2 into two variables
```{r}
dfb = dfa %>% separate(col=term2,into=c("semester","year"), sep=" ")
```

### Clean and subset the data
we get rid of the following variables:

* acad_org
* current_status
* instructor
* course_id

Filtering to only include:

* Math courses
* Lectures

We remove duplicate rows using distinct()
```{r}
dfc = dfb %>% select(-acad_org,-current_status,-instructor,-course_id) %>% filter(subject=="MA") %>% filter(catalog%in% c("131","132"), semester %in% c("Fall","Spring")) %>% filter(component=="LEC") %>% distinct()

```

### Summarising the data
Group by:

* class
* semester
* year

Sum up enrollment across many sections by summarizing by sum()


```{r}
dfd = dfc %>% group_by(title,semester,year) %>% summarise(total_enrolment_amt =sum(tot_enrl))
```

