---
title: "Class enrollment Data"
author: "Joe Tudor"
date: "18 Sep 2020"
output: html_notebook
---

Notebook Created for the purpose of learning different tools for analysis in R whilst looking for relations in the class enrollment data at Clarkson University. 

Prepare workspace:

## Load packages

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(skimr)
library(summarytools)
```

## Read the data
Creates first data frame using clean_names() in order to start tidying up the data and making the names more uniform and easy to use in later functions.
```{r}
dfa= read_excel("CU_ACAD_DEPT_ENRL.xlsx") %>%
clean_names()
```

Skimming throught the data set to get a rough look at the data before analysis.
```{r}
skim(dfa)
```
```{r}
C=dfa %>% group_by(subject) %>% skim()
```




```{r}
dfa %>% dfSummary()
```

Filtering for just Spring 2019.
```{r }
spring19 <- dfa %>% filter(term2=="Spring 2019") %>% dfSummary() %>% view()
skim(spring19)
```


```{r }
dfa %>% dfSummary() %>% view()
```


```{r message=FALSE, warning=FALSE, results='asis'}
print(dfSummary(dfa, graph.magnif = 0.75), method = 'render')
```


